---
title: "comparison of drug and achilles targets"
author: "Lennard"
date: "24/6/2021"
output: html_document
---


```{r}
load("../2021-topic-01-team-03/dataset/prism_datasets.rda")
load("../2021-topic-01-team-03/dataset/cellline_datasets.rda")
```

# Guidelines

prism: effect of the treatment (columns) on cell growth of cell lines (rows).For example, BRD-A00077618-236-07-6::0.0024414::HTS002 indicates treatment with the drug BRD-A00077618-236-07-6, with a dose of 2.4 nM, in assay HTS002. Values represent log fold changes with respect to a control DMSO experiment.

prism.treat:information about the treatment (row) and the drug used

prism.cl: cell lines' info

prism.exp: gene TPM (transcripts per million) values. Indicate over/underexpression. Rows = cell lines, columns = genes

prism.cnv: gene copy number (CN) values. In absolute terms, CN = 2, since there are two alleles per genes. In cancer, genes might be amplified CN > 2 or deleted CN < 2

prism.snv: observed mutations in a sample. The isDeleterious flag specifies if mutation has a functional effect or not. First column is genes. There is a column
(DepMap_ID) with corresponding cell lines

prism.achilles: gene KO scores, a measure of how essential/important is a particular gene for the cell survival

IDEA: use same indications as in PRISM paper to start filtering for relevant genes, cell lines? 


# Cleaning

## Selecting brain cancer specific cell lines in all data sets

```{r Selection of cell lines}
brain_cancer_cl=subset.data.frame(prism.cl, disease == 	"Brain Cancer") #identify the brain cancer cell lines' names

names=brain_cancer_cl[,1] #saving the names as names

#select wanted cell lines from all datasets
brain_cancer=prism[names,]
brain_cancer_achilles=prism.achilles[names,]
brain_cancer_cnv=prism.cnv[names,]
brain_cancer_exp=prism.exp[names,]
```

## Working on NAs

```{r Working on NAs}
# replacing NAs 
lines=row.names(brain_cancer)
row.names(brain_cancer_achilles) = lines

# NA removal 
brain_cancer_achilles_no_NA=na.omit(brain_cancer_achilles) #remove the entire cell lines from everywhere without KO scores

# NA substitution for mean in brain_cancer
for(i in 1:ncol(brain_cancer)){
  brain_cancer[is.na(brain_cancer[,i]), i] <- mean(brain_cancer[,i], na.rm = TRUE)
}
brain_cancer=brain_cancer[ , colSums(is.na(brain_cancer)) == 0]
```

## Removal of useless columns 

```{r}
brain_cancer_cl=brain_cancer_cl[,-c(4,6,7,8,9,11,12,15,17,19,23,24,26,28,29,30,31,32,33)]
```

## Dividing cell lines into subtypes

```{r}
medulloblastoma=brain_cancer_cl$DepMap_ID[grep("Medulloblastoma", brain_cancer_cl$disease_subtype)]#we obtain the names of the cell lines that are classified as medulloblastoma

glioblastoma=brain_cancer_cl$DepMap_ID[grep("Glioblastoma", brain_cancer_cl$disease_subtype)]

glioma=brain_cancer_cl$DepMap_ID[grep("Glioma", brain_cancer_cl$disease_subtype)]

astrocytoma=brain_cancer_cl$DepMap_ID[grep("Astrocytoma", brain_cancer_cl$disease_subtype)]

#making sure we didnt forget anything 

disease_subtypes=c(medulloblastoma,glioblastoma,glioma,astrocytoma)
length(disease_subtypes)

# Converting them into dataframes [IS THIS RELEVANT?]
glioblastoma_bc <- brain_cancer[glioblastoma,]
medulloblastoma_bc <- brain_cancer[medulloblastoma,]
glioma_bc <- brain_cancer[glioma,]
astrocytoma_bc <- brain_cancer[astrocytoma,]
```


# Question 1: How can we distinguish the most effective drugs?

## Dividing drugs in doses groups

```{r}
dd1=grep("::0.00061034::", colnames(brain_cancer))
dd2=grep("::0.0024414::", colnames(brain_cancer))
dd3=grep("::0.00976562::", colnames(brain_cancer))
dd4=grep("::0.0390625::", colnames(brain_cancer))
dd5=grep("::0.15625::", colnames(brain_cancer))
dd6=grep("::0.625::", colnames(brain_cancer))
dd7=grep("::10::", colnames(brain_cancer))
dd8=grep("::2.5::", colnames(brain_cancer))
#positions of the respective doses in the dataframe 
doses=c(dd1,dd2,dd3,dd4,dd5,dd6,dd7,dd8)



doses_for_list=list("::0.00061034::","::0.0024414::","::0.00976562::","::0.0390625::","::0.15625::","::0.625::","::10::","::2.5::")
select_doses=function(x){
  brain_cancer[,grep(x, colnames(brain_cancer))]
}
doses_subsets_<-lapply(doses_for_list,select_doses)



d1=brain_cancer[,dd1]
d2=brain_cancer[,dd2]
d3=brain_cancer[,dd3]
d4=brain_cancer[,dd4]
d5=brain_cancer[,dd5]
d6=brain_cancer[,dd6]
d7=brain_cancer[,dd7]
d8=brain_cancer[,dd8]

#these are the standard doses summed up into seperate dataframes
sum(length(dd1), length(dd2), length(dd3), length(dd4),length(dd5), length(dd6),length(dd7),length(dd8)) #ups something missing

deviation=brain_cancer[,-doses] #lets see what we missed-> these are the doses that are not exact and have some kind of a deviation

questionable_drugs=prism.treat[colnames(deviation),]#we dont know if we can use these drugs, may be their doses deviate too far from the standard
```

## Working on drugs that do not fit the standard doses

```{r}
#Stefan's suggestion: assigning the outliers to closest standard doses

#Calculation of all differences between questionable drugs and standard doses
std_doses <- c(0.00061034,0.0024414,0.00976562,0.0390625,0.15625,0.625,10,2.5) #vector with standard dosages

#Picking dosages from dosages that have to be assigned to standard dosages
doses_to_assign=as.numeric(questionable_drugs$dose)

#Creating data frame for the differences between standard dosages and questionable drugs
extra_drugs <-data.frame(matrix(nrow = 8,ncol = length(doses_to_assign)),row.names = std_doses)
colnames(extra_drugs)=rownames(questionable_drugs)

for (i in 1:length(doses_to_assign)){
    extra_drugs[,i]=abs(std_doses - doses_to_assign[i])
}

#Assigning the drugs to the closest standard dosage 
to_which_dose=apply(extra_drugs, 2, which.min) #function of which.min: gives row number of min value of column. In comparison just apply: gives min difference

d1_extra=which(to_which_dose==1)
d2_extra=which(to_which_dose==2)
d3_extra=which(to_which_dose==3)
d4_extra=which(to_which_dose==4)
d5_extra=which(to_which_dose==5)
d6_extra=which(to_which_dose==6)
d7_extra=which(to_which_dose==7)
d8_extra=which(to_which_dose==8)



doses_for_list_2<-list(1,2,3,4,5,6,7,8)
select_doses_2=function(x){
  brain_cancer[,grep(x, to_which_dose)]
}
doses_subsets_2<-lapply(doses_for_list_2,select_doses_2)

add_extra_doses_list=function(x){
  cbind(doses_subsets_2[[x]], doses_subsets_[[x]])
}
doses_subsets_<-lapply(doses_for_list_2,add_extra_doses_list)



#just a quick check how many doses there are in total
not_standard_doses=c(d1_extra,d2_extra,d3_extra,d4_extra,d5_extra,d6_extra,d7_extra,d8_extra)
length(not_standard_doses)#in comparison with dimension from questionable_drugs right!

#Adding new assigned dosages to small data frames of standard dosages 
add_extra_doses=function(x,y){
  cbind(brain_cancer[,unique(grep(paste(colnames(extra_drugs[,x]),collapse="|"), colnames(brain_cancer)))], y)
  }

d1=add_extra_doses(d1_extra,d1)
d2=add_extra_doses(d2_extra,d2)
d3=add_extra_doses(d3_extra,d3)
d4=add_extra_doses(d4_extra,d4)
d5=add_extra_doses(d5_extra,d5)
d6=add_extra_doses(d6_extra,d6)
d7=add_extra_doses(d7_extra,d7)
d8=add_extra_doses(d8_extra,d8)



#Resumé: new data sets for each dosage with assigned outliers to closest deviation

ncol(d2)+ncol(d1)+ncol(d3)+ncol(d4)+ncol(d5)+ncol(d6)+ncol(d7)+ncol(d8)
ncol(brain_cancer)
```


## Filter effective drugs in each dose group

```{r}
#Lets say for now that our threshold is 0,3 so we can see how many drugs meet this criteria in each dose group
effective_drugs<-function(x){
  effective_x<-x[,which(apply(x,2,mean)<=0.2)]
  return(effective_x)
}

effective_d1=effective_drugs(d1)
effective_d2=effective_drugs(d2)
effective_d3=effective_drugs(d3)
effective_d4=effective_drugs(d4)
effective_d5=effective_drugs(d5)
effective_d6=effective_drugs(d6)
effective_d7=effective_drugs(d7)
effective_d8=effective_drugs(d8)


setdiff(colnames(d1), colnames(doses_subsets_[[1]])) #something is wrong
doses_subsets <- list(d1 = d1, d2 = d2, d3 = d3, d4 = d4, d5 = d5, d6 = d6, d7 = d7, d8 = d8)
all_effective_drugs<- lapply(doses_subsets, effective_drugs)



ncol(all_effective_drugs[[1]])+ncol(all_effective_drugs[[2]])+ncol(all_effective_drugs[[3]])+ncol(all_effective_drugs[[4]])+ncol(all_effective_drugs[[5]])+ncol(all_effective_drugs[[6]])+ncol(all_effective_drugs[[7]])+ncol(all_effective_drugs[[8]])

length(all_effective_drugs)
length(which(apply(brain_cancer,2,mean)<=0.2)) #this is the amount of drugs that we keep in total. We compare these 2 length to make sure that that we do not select the same drug twice. If it was the case the first length would be bigger than the second one
```

##Select drugs that are effective in all doses

```{r}
save_new_colnames<-function(x){
  a<-gsub(pattern = "::.*", replacement = "", colnames(x))
  colnames(x)<-a
  return(x)
}

effective_d1<-save_new_colnames(effective_d1)
effective_d2<-save_new_colnames(effective_d2)
effective_d3<-save_new_colnames(effective_d3)
effective_d4<-save_new_colnames(effective_d4)
effective_d5<-save_new_colnames(effective_d5)
effective_d6<-save_new_colnames(effective_d6)
effective_d7<-save_new_colnames(effective_d7)
effective_d8<-save_new_colnames(effective_d8)



all_effective_drugs<-lapply(all_effective_drugs, save_new_colnames)
## remove everything but the drug information from colnames and make them new colnames of the dataframes


  
common_drugs<-function(x,y){
  different<-setdiff(colnames(x), colnames(y))
  
  different<-unique(grep(paste(different,collapse="|"), colnames(x)))
  
  colnames(x[,-different])
}
#function returns the names of the common drugs between 2 dataframes

d1_d2=common_drugs(effective_d1,effective_d2)
d2_d1=common_drugs(effective_d2,effective_d1)
#both are equal, so the function works correctly

d1_d3=common_drugs(effective_d1,effective_d3)
d1_d4=common_drugs(effective_d1,effective_d4)
d1_d5=common_drugs(effective_d1,effective_d5)
d1_d6=common_drugs(effective_d1,effective_d6)
d1_d7=common_drugs(effective_d1,effective_d7)
d1_d8=common_drugs(effective_d1,effective_d8)

#which drugs are present in all 8 doses?

present_in_all_doses<-function(a,b,c,d,e,f,g){
 bb<-setdiff(a, b)
 a<-a [! a %in% bb]
 
 cc<-setdiff(a, c)
 a<-a [! a %in% cc]
 
 dd<-setdiff(a, d)
 a<-a [! a %in% dd]
 
 ee<-setdiff(a, e)
 a<-a [! a %in% ee]
 
 ff<-setdiff(a, f)
 a<-a [! a %in% ff]
 
 gg<-setdiff(a, g)
 a<-a [! a %in% gg]
 
 return(a)
}

effective_in_all_doses=present_in_all_doses(d1_d2,d1_d3,d1_d4,d1_d5,d1_d6,d1_d7,d1_d8)
#51 drugs are effective in all doses
```

# Question 2: What are the targets of the effective drugs?

##What are the targets of the effective drugs?

```{r}
brain_cancer_treat=prism.treat[colnames(brain_cancer),] #selecting only the drugs we previously filtered out from brain_cancer
brain_cancer_treat=brain_cancer_treat[!duplicated(brain_cancer_treat$broad_id),]
rownames(brain_cancer_treat)<-brain_cancer_treat[,1]
effective_bc_treat=brain_cancer_treat[effective_in_all_doses,-c(1,2,3,4,10)]
#summary(effective_bc_treat$target)
#summary(effective_bc_treat$moa)
#expression of these genes would be interesting to look into our cell lines
```

#Target of effective drugs
```{r}
brain_cancer.effective_targets <- effective_bc_treat$target

which(is.na(effective_bc_treat$target))

## da fehlen halt bei diesen 8 Drogen die targets, die könnte man sonst ggf. rausnehmen 
```


##Selecting brain cancer specific cell lines in all data sets
```{r}
brain_cancer_cl=subset.data.frame(prism.cl, disease == 	"Brain Cancer") #identify the brain cancer cell lines' names

names=brain_cancer_cl[,1] #saving the names as names

#select wanted cell lines from all datasets
brain_cancer=prism[names,]
brain_cancer_achilles=prism.achilles[names,]

## hier ist irgendwas schief gegangen, ab da sind super viele NA's drin 

brain_cancer_cnv=prism.cnv[names,]
brain_cancer_exp=prism.exp[names,]
#brain_cancer_snv=prism.snv[names,] #does not make sense no cell line info, other type of filtering required
#brain_cancer_treat=prism.treat[names,] #does not make sense no cell line info, other type of filtering required
```

# Selecting achilles targets
```{r}
#brain_cancer_achilles.mean <- apply(brain_cancer_achilles,2,mean)
#hist(brain_cancer_achilles.mean, breaks = 200)
#brain_cancer_achilles.var <- apply(brain_cancer_achilles,2,var)#variance
#hist(brain_cancer_achilles.var, breaks = 200, xlim = c(0,0.4))
#brain_cancer_achilles.relevant <- colnames(brain_cancer_achilles)[brain_cancer_achilles.mean < -0.2]
```

# Compare drug and achilles targets
```{r}
brain_cancer.effective_targets <- as.character(brain_cancer.effective_targets) # convert to character

x <- 1
brain_cancer.effective_targets.solo <- list() #create list

for(i in brain_cancer.effective_targets) #split multiple targets
  {
  
 brain_cancer.effective_targets.solo[[x]] <- strsplit(i, split = ", ")
 x <- x+1
 }
brain_cancer.effective_targets.solo <- unlist(brain_cancer.effective_targets.solo) #list to vector

brain_cancer.effective_targets.solo <- unique(brain_cancer.effective_targets.solo) # remove multiple targets

brain_cancer.effective_targets.solo <- brain_cancer.effective_targets.solo[-1] # remove NA

# achilles targets
brain_cancer_achilles.targets <- brain_cancer_achilles[,which(colnames(brain_cancer_achilles) %in% brain_cancer.effective_targets.solo)] # selecting target genes in achilles
brain_cancer.effective_targets.solo[-which(brain_cancer.effective_targets.solo %in% colnames(brain_cancer_achilles))] # missing targets

hist(unlist(brain_cancer_achilles), breaks = 50)
hist(unlist(brain_cancer_achilles.targets), breaks = 50)
summary(unlist(brain_cancer_achilles))
summary(unlist(brain_cancer_achilles.targets))
qqplot(unlist(brain_cancer_achilles), unlist(brain_cancer_achilles.targets))
abline(a = 0, b = 1)

# achilles variance analysis
brain_cancer_achilles.var <- apply(brain_cancer_achilles,2,function(x){var(x, na.rm = TRUE)})#variance
brain_cancer_achilles.targets.var <- apply(brain_cancer_achilles.targets,2, function(x){var(x, na.rm = TRUE)})
hist(unlist(brain_cancer_achilles.var), breaks = 500, xlim = c(0,0.2))
hist(unlist(brain_cancer_achilles.targets.var), breaks = 40, xlim = c(0,0.2))
qqplot(unlist(brain_cancer_achilles.var), unlist(brain_cancer_achilles.targets.var), xlim = c(0,0.2))
abline(a = 0, b = 1)


# exp targets
brain_cancer_exp.targets <- brain_cancer_exp[,which(colnames(brain_cancer_exp) %in% brain_cancer.effective_targets.solo)] # selecting target genes in exp
brain_cancer.effective_targets.solo[-which(brain_cancer.effective_targets.solo %in% colnames(brain_cancer_exp))] # missing targets

hist(unlist(brain_cancer_exp), breaks = 50)
hist(unlist(brain_cancer_exp.targets), breaks = 50)
summary(unlist(brain_cancer_exp))
summary(unlist(brain_cancer_exp.targets))
qqplot(unlist(brain_cancer_exp), unlist(brain_cancer_exp.targets))
abline(a = 0, b = 1)

# exp variance analysis
brain_cancer_exp.var <- apply(brain_cancer_exp,2,function(x){var(x, na.rm = TRUE)})#variance
brain_cancer_exp.targets.var <- apply(brain_cancer_exp.targets,2, function(x){var(x, na.rm = TRUE)})
hist(unlist(brain_cancer_exp.var), breaks = 500)
hist(unlist(brain_cancer_exp.targets.var), breaks = 40)
qqplot(unlist(brain_cancer_exp.var), unlist(brain_cancer_exp.targets.var), xlim = c(0,8))
abline(a = 0, b = 1)


# cnv targets
brain_cancer_cnv.targets <- brain_cancer_cnv[,which(colnames(brain_cancer_cnv) %in% brain_cancer.effective_targets.solo)] # selecting target genes in cnvVie
brain_cancer.effective_targets.solo[-which(brain_cancer.effective_targets.solo %in% colnames(brain_cancer_cnv))] # missing targets

hist(unlist(brain_cancer_cnv), breaks = 50)
hist(unlist(brain_cancer_cnv.targets), breaks = 50)
summary(unlist(brain_cancer_cnv))
summary(unlist(brain_cancer_cnv.targets))
qqplot(unlist(brain_cancer_cnv), unlist(brain_cancer_cnv.targets))
abline(a = 0, b = 1)

# cnv variance analysis
brain_cancer_cnv.var <- apply(brain_cancer_cnv,2,function(x){var(x, na.rm = TRUE)})#variance
brain_cancer_cnv.targets.var <- apply(brain_cancer_cnv.targets,2, function(x){var(x, na.rm = TRUE)})
hist(unlist(brain_cancer_cnv.var), breaks = 500)
hist(unlist(brain_cancer_cnv.targets.var), breaks = 40)
qqplot(unlist(brain_cancer_cnv.var), unlist(brain_cancer_cnv.targets.var), xlim = c(0,0.1), ylim = c(0,0.1))
abline(a = 0, b = 1)
```

