---
title: "Bioinfo Projekt"
author: "Ilya"
date: "5/1/2021"
output: html_document
---

```{r}
load("C:/Users/schni/OneDrive/Documents/Germany/4. Semester/Projekt Bioinfo/prism_datasets.rda")
load("C:/Users/schni/OneDrive/Documents/Germany/4. Semester/Projekt Bioinfo/cellline_datasets.rda")
```

#Guidelines

prism: effect of the treatment (columns) on cell growth of cell lines (rows).For example, BRD-A00077618-236-07-6::0.0024414::HTS002 indicates treatment with the drug BRD-A00077618-236-07-6, with a dose of 2.4 nM, in assay HTS002. Values represent log fold changes with respect to a control DMSO experiment.

prism.treat:information about the treatment (row) and the drug used

prism.cl: cell lines' info

prism.exp: gene TPM (transcripts per million) values. Indicate over/underexpression. Rows = cell lines, columns = genes

prism.cnv: gene copy number (CN) values. In absolute terms, CN = 2, since there are two alleles per genes. In cancer, genes might be amplified CN > 2 or deleted CN < 2

prism.snv: observed mutations in a sample. The isDeleterious flag specifies if mutation has a functional effect or not.

prism.achilles: gene KO scores, a measure of how essential/important is a particular gene for the cell survival

IDEA: use same indications as in PRISM paper to start filtering for relevant genes, cell lines? 

#Cleaning

##Selecting brain cancer specific cell lines in all data sets
```{r}
brain_cancer_cl=subset.data.frame(prism.cl, disease == 	"Brain Cancer") #identify the brain cancer cell lines' names

names=brain_cancer_cl[,1] #saving the names as names

#select wanted cell lines from all datasets
brain_cancer=prism[names,]
brain_cancer_achilles=prism.achilles[names,]
brain_cancer_cnv=prism.cnv[names,]
brain_cancer_exp=prism.exp[names,]
brain_cancer_snv=prism.snv[names,] #didnt work
brain_cancer_treat=prism.treat[names,] #didnt work
```

##NA removal
```{r}
brain_cancer_achilles=na.o mit(brain_cancer_achilles) #remove the entire cell lines from everywhere without KO scores?

```


